---
title: "Homework assignment 1"
author: "Nour Al Afif"
execute:
  echo: true
  warning: false
format: pdf
---

#### Question 1 & 2: Loading Library

```{r}
library(babynames)
library(tidyverse)
glimpse(babynames)
```

It contains 5 variables and 1,924,665 observations.

#### Question 3 Data dictionary

Column 1: year- indicates the year that the person was born Column 2: sex - the sex of the child given the name Column 3: name - the name being recorded Column 4: n - the number of times the name was used in the given year (Column 1) Column 5: prop - the proportion of people of a specific sex with the given name in a certain year, calculated by dividing n (column 4) by the total number of observations for any given year.

#### Question 4 Range of years in file

```{r}
range(babynames$year)
```

#### Question 5 & 6 Data frame without n

```{r}
babynames_props <- select(babynames, 1:3, 5)
```

n is a very subjective measure and it is hard to know how common a name is based on just that, because popularity in any given year would be related to the number of people born that year.

#### Question 7 Most popular names in the second and third millenia

```{r}
pop_sec_mil <- babynames_props |>
  filter(year %in% c(1900:2000))|>
  group_by(name, sex)|>
  summarize(prop= sum(prop))|>
  group_by(sex)|>
  filter(prop == max(prop, na.rm=TRUE))
pop_sec_mil

pop_third_mil <- babynames_props |>
  filter(year >= 2000)|>
  group_by(name, sex)|>
  summarize(prop= sum(prop))|>
  group_by(sex)|>
  filter(prop == max(prop, na.rm=TRUE))

pop_third_mil
```

#### Question 8
```{r}
pop_qvx <- babynames_props |>
  filter(year %in% c(2000:2012), substr(name,1,1) %in% c("Q", "V", "X"))|>
  group_by(name)|>
  summarize(prop= sum(prop))|>
  group_by(substr(name,1,1))|>
  filter(prop == max(prop, na.rm=TRUE))

pop_qvx
```

#### Question 9 and 10 
```{r}
decades <- babynames |>
  mutate(decade = paste(substr(year,1,3), 0, "s", sep = ""))

sex_data <- decades|>
  group_by(decade, sex)|>
  summarize(median = median(n), mean = mean(n))
sex_data
  
```

#### Question 11
##### Most popular decade
```{r}
names <- decades |>
  filter(name %in% c("Nour", "Michael", "Jack", "Scott"))|>
  group_by(name)

names|>
  group_by(name, decade)|>
  summarize(prop = sum(prop, na.rm=TRUE))|>
  filter(prop == max(prop, na.rm = TRUE))
```

##### Most popular year
```{r}
names|>
  filter(prop == max(prop, na.rm = TRUE))

```
