---
title: "KINESIOL 736 -- Assignment 2"
author: "Nour Al Afif"
format: pdf
---
### Part B
#### 1. Reading in the file
```{r}
#| warning: false
#| message: false
library(tidyverse)
lotr_data <- read_csv("lord-of-the-rings-trilogy.csv")
```
###### 2. List all the ways this dataset is not tidy.
- There isn't consistent naming conventions, with some of the races/sexes being capitalized in some instances but not the others.
- Each variable doesn't have its own column and each row doesn't have its own observation.

###### 3. How many columns and rows would you have if this dataset was tidy?
The dataset would have 3 columns and 18 rows, because there are 3 movies x 3 races x 2 sexes

###### 4. What would the column names be in tidy format?
The column names would be: movie, sex, race, word_count

### Part B
##### 1. Tidying the dataset
```{r}
#| message: false
lotr_tidy <- lotr_data |>
  pivot_longer(elf_female:Man_male,
               names_to = "race_sex",
               values_to = "word_count")|>
  mutate(race_sex = tolower(race_sex))|>
  tidyr::separate(race_sex, into = c("race", "sex"), convert = TRUE)
```

##### 2. What is the total number of words spoken by:
a) male hobbits
```{r}
#| message: false
lotr_tidy |> group_by(race, sex)|>
  filter(sex == "male", race == "hobbit")|>
  summarize(word_count = sum(word_count))
```
b) female elves
```{r}
#| message: false
lotr_tidy |> group_by(race, sex)|>
  filter(sex == "female", race == "elf")|>
  summarize(word_count = sum(word_count))
```
c) male elves
```{r}
#| message: false
lotr_tidy |> group_by(race, sex)|>
  filter(sex == "male", race == "elf")|>
  summarize(word_count = sum(word_count))
```

##### 3. Is the number of spoken words in a movie dominated by a single race?
```{r}
lotr_tidy |> group_by(movie, race)|>
  summarize(word_count = sum(word_count))|>
  filter(word_count == max(word_count))|>
  ungroup()
```
Yes, man is dominant in the return of the king and two towers while hobbits are dominant in the fellowship of the ring.

##### 4. Does the dominant race depend on the movie?
Yes it does, the fellowship of the ring has most words spoken by hobbits while the other two movies are dominated by man.

